% -*-latex-*-

%% Defines the cover material, including the title page, copyright page,
%% signature page, abstract, and acknowledgements.

\newgeometry{left=3.5cm,bottom=0.1cm}

\title{Multiscale analysis of infectious diseases:\\
       Integrating omics and clinical informatics data into patient care}
\fancytitle{
  \def\baselinestretch{1.1}\Huge
  \textcolor{AccentColor}{\smallcapsspacing{\MakeUppercase{Multiscale analysis}}}\\
  \textit{of}\\
  \textcolor{AccentColor}{\smallcapsspacing{\MakeUppercase{infectious diseases}}}\\
  \vspace{1cm}
  \LARGE
  \smallcapsspacing{\MakeLowercase{\scshape Integrating omics and clinical informatics data}}\\
  \def\baselinestretch{0.5}
  \smallcapsspacing{\MakeLowercase{\scshape into patient care}}
}
\author{Theodore Robertson Pak}

\degree{Doctor of Philosophy}
\program{Biomedical Sciences Doctoral Program}
\department{Graduate School of Biomedical Sciences}
\school{Icahn School of Medicine at Mount Sinai}

\degreemonth{April}
\degreeyear{2017}
\thesisdate{April 27, 2017}

\supervisor{Andrew Kasarskis, PhD}
\dean{Marta Filizola, PhD}{Dean of the Graduate School of Biomedical Sciences}

\committeemembers{
  James Iatridis, PhD (chair)\\
  Adolfo García-Sastre, PhD\\
  Joel Dudley, PhD\\
  Jonathan Karr, PhD\\
  Jun Zhu, PhD\\
  Bo Shopsin, MD, PhD (outside reviewer)
}

% Make the titlepage based on the above information.  If you need
% something special and can't use the standard form, you can specify
% the exact text of the titlepage yourself.  Put it in a titlepage
% environment and leave blank lines where you want vertical space.
% The spaces will be adjusted to fill the entire page.  The dotted
% lines for the signatures are made with the \signature command.
\maketitle

% standard copyright page required by ISMMS formatting
\copyrightpage

% standard signature page required by ISMMS formatting
\signaturepage

\restoregeometry

%% Either \input (*not* \include) your abstract file, or you can put
%% the text of the abstract directly between the \begin{abstractpage} and
%% \end{abstractpage} commands.

\begin{abstractpage}
\input{abstract}
\end{abstractpage}

\cleardoublepage


\chapter*{Acknowledgments}
Suspendisse consequat, nisl sed feugiat dignissim, enim mi posuere lacus, ut dictum dolor dolor quis augue. Curabitur porta urna nec felis accumsan, sed commodo massa auctor. Vivamus interdum est ac efficitur malesuada. Quisque odio ligula, finibus at pharetra et, dictum at eros. Aenean a lorem in urna volutpat maximus vestibulum eget augue. Nullam elementum ipsum quis mauris luctus, ac cursus metus aliquam. Curabitur sed odio imperdiet nunc pulvinar venenatis nec vehicula neque. In ac sollicitudin lorem. Vestibulum tempus ante nisi, laoreet accumsan felis dictum vitae.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% -*-latex-*-